---
title: ASMAS SS2024 - Exercise 1
author: "Peter von Rohr"
date: 2024-02-19
filters: 
  - webr
params:
  doctype:
    label: 'Document Type'
    value: 'solution'
    choices: ['exercise', 'solution', 'notebook']
  number:
    label: 'Number'
    value: 1
    choices: c(1:15)  
  isonline:
    label: 'Online (y/n)'
    value: TRUE
    choices: [TRUE, FALSE]
---


## Problem 1: Data Collection and First Inspection
Measure width of your left hand in mm and your height in cm and enter that data into a table. From that table create a simple plot using the function `plot()`. As a demo, we are first running that with random numbers. After that run the same with the collected data

```{r}
#| label: demoplot
set.seed(1902)
n_nr_obs <- 15
vec_width <- rnorm(n_nr_obs, mean = 73, sd = 2.5)
vec_height <- rnorm(n_nr_obs, mean = 175, sd = 7.9)
plot(vec_width, vec_height)
```

Do the same with the collected data

```{webr-r}
# enter data

# plot

```



Order both variables (width and height) in the data according to the width of the left hand, then repeat the plot

```{r}
vec_order <- order(vec_width)
vec_ord_width <- vec_width[vec_order]
vec_ord_height <- vec_height[vec_order]
plot(vec_ord_width, vec_ord_height)
```

Do the ordered version of the plot

```{web-r}

```


## Problem 2: Line Fitting
* Based on the plotted points above what would be a good description of the relationship of width and height?
* How well would a straight line describe the relationship?
* Why do we want to use a straight line? 
* What is the meaning of a linear relationship?
* Try to fit different straight lines throught the plotted points by trial and error. 
* Try to make statements about how well a given straight line fits the points. 


