---
title: "Interactions"
author: "Peter von Rohr"
date: "2023-04-03"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Definition

* Effect of given predictor variable depends on level or value of other predictor variable
* Examples:
    + Regression of `Body Weight` on `Breast Circumference` is different for different breeds
    + Effect of `Breed` on `Body Weight` is different for different male and female animals


## Types of Interactions

* continuous by continuous
* continuous by categorical
* categorical by categorical


## Continuous by Categorical

* In a model, expected value of response depends on regression variable plus discrete factor
* Example: Regression of `Body Weight` on `Breast Circumference` plus the factor `Breed`
* Interaction is present, if regression of `Body Weight` on `Breast Circumference` is different for different breeds


## Interaction Plot

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
s_data_dir <- file.path(here::here(), "docs", "data")
s_flem_path <- file.path(s_data_dir, "asm_bw_flem.csv")
tbl_flem <- readr::read_csv(file = s_flem_path)
tbl_flem_bw_bc_br <- tbl_flem %>%
  select(Animal, `Breast Circumference`, `Body Weight`, Breed)
tbl_flem_bw_bc_br <- tbl_flem_bw_bc_br %>% arrange(Breed)
formula_bw_bc <- `Body Weight` ~ `Breast Circumference`
tbl_flem_bw_bc_br_an <- tbl_flem_bw_bc_br %>%
  filter(Breed == "Angus")
lm_bw_bc_br_an <- lm(formula = formula_bw_bc, data = tbl_flem_bw_bc_br_an)
#lm_bw_bc_br_an$coefficients[["(Intercept)"]]
#lm_bw_bc_br_an$coefficients[["`Breast Circumference`"]]
tbl_flem_bw_bc_br_li <- tbl_flem_bw_bc_br %>%
  filter(Breed == "Limousin")
lm_bw_bc_br_li <- lm(formula = formula_bw_bc, data = tbl_flem_bw_bc_br_li)
tbl_flem_bw_bc_br_si <- tbl_flem_bw_bc_br %>%
  filter(Breed == "Simmental")
lm_bw_bc_br_si <- lm(formula = formula_bw_bc, data = tbl_flem_bw_bc_br_si)

# plots
library(ggplot2)
p <- ggplot(data = tbl_flem_bw_bc_br, aes(x=`Breast Circumference`, y = `Body Weight`, group = Breed)) + 
  geom_point(aes(color = Breed)) + 
  geom_abline(intercept = lm_bw_bc_br_an$coefficients[["(Intercept)"]],
              slope = lm_bw_bc_br_an$coefficients[["`Breast Circumference`"]],
              color = "red") + 
  geom_abline(intercept = lm_bw_bc_br_li$coefficients[["(Intercept)"]],
              slope = lm_bw_bc_br_li$coefficients[["`Breast Circumference`"]],
              color = "green") + 
  geom_abline(intercept = lm_bw_bc_br_si$coefficients[["(Intercept)"]],
              slope = lm_bw_bc_br_si$coefficients[["`Breast Circumference`"]],
              color = "blue")
p
```


## Interaction Model

* Start with model without interactions

$$y_i = b_0 + b_1 \times BC_i + b_2 \times BrLi_i + b_3 \times BrSi_i + e_i$$

* Assume linear relationship of $b_1$ with `Breed`

$$b_1 = a + b_4 \times BrLi + b_5 \times BrSi$$

* Insert

$$y_i = b_0 + (a + b_4 \times BrLi + b_5 \times BrSi) \times BC_i + b_2 \times BrLi_i + b_3 \times BrSi_i + e_i$$

* Simplify

\begin{align}
y_i &= b_0 + a \times BC_i + b_2 \times BrLi_i + b_3 \times BrSi_i \notag \\
    &  + b_4 \times BrLi \times BC_i + b_5 \times BrSi \times BC_i + e_i \notag
\end{align}


## Continuous by Continuous

* Similar to continuous by categorical
* No interaction

\begin{equation}
y_0 = b_0 + b_1 \times BC_i + b_2 \times HE_i + e_i \notag
\end{equation}

* Interaction by dependence of one regression coefficient on other coefficient

$$b_1 = b_3 + b_4 \times HE_i $$

$$y_0 = b_0 + b_2 \times HE_i + b_3 \times BC_i + b_4 \times HE_i \times BC_i + e_i$$


## Categorical by Categorical 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
tbl_flem_bw_br_sex <- tbl_flem_bw_bc_br %>%
  select(Animal, `Body Weight`, Breed)
tbl_flem_bw_br_sex <- tbl_flem_bw_br_sex %>% 
  bind_cols(tibble::tibble(Sex = c("M", rep("F",2),
                                   rep("F", 2), "M", "F",
                                   "F", "M","F")))
library(ggplot2)
p <- ggplot(data = tbl_flem_bw_br_sex, 
            aes(x = Breed, 
                y = `Body Weight`,
                group = Sex)) + 
  geom_point(aes(color = Sex))
p

```


## Model Matrix

\tiny

```{r}
lm_bw_br_sex <- lm(`Body Weight` ~ Breed * Sex, data = tbl_flem_bw_br_sex)
model.matrix(lm_bw_br_sex)
```


## Summary

\tiny

```{r}
summary(lm_bw_br_sex)
```

