---
title: 'ASM Exam 2024: Problem 2'
author: "Peter von Rohr"
date: "2024-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Background
Include breed as additional fixed effects into the dataset. Use parameter from Problem 1 ...

```{r}
set.seed(2405)
n_nr_obs <- 20
# dim
n_dim_start <- 20
n_dim_end <- 50
# citrate
n_cit_max <- 9.2
n_cit_min <- 8.5
n_cit_mean <- mean(c(n_cit_max, n_cit_min))
n_cit_sd <- 0.24
n_reg_coef <- (n_cit_min - n_cit_max) / (n_dim_end - n_dim_start)
n_reg_intc <- 9.5

```


## Data Simulation
The DIM values are sampled from a uniform distribution determined by the DIM-range.

```{r}
vec_dim <- round(runif(n_nr_obs, min = n_dim_start, max = n_dim_end), digits = 0)
vec_dim <- vec_dim[order(vec_dim)]
vec_dim
```

The vector of breeds is sampled randomly

```{r}
vec_breed_eff <- c(0, 0.46)
vec_breed_name <- c("BS", "HO")
vec_breed <- sample(vec_breed_name, n_nr_obs, replace = T)
vec_breed
```

The data is generated using breed as factor

```{r}
vec_breed_fct <- as.factor(vec_breed)
mat_X_breed <- model.matrix(y ~ breed, data = data.frame(y = rep(1, n_nr_obs), breed = vec_breed_fct))
attr(mat_X_breed, "assign") <- NULL
attr(mat_X_breed, "contrasts") <- NULL
dimnames(mat_X_breed) <- NULL
mat_X_breed
```



Generate the data

```{r}
vec_cit <- n_reg_intc + n_reg_coef * vec_dim + crossprod(t(mat_X_breed), vec_breed_eff) + rnorm(n_nr_obs, mean = 0, sd = n_cit_sd)
vec_cit_round <- round(vec_cit, digits = 2)
tbl_cit_dim_br <- tibble::tibble(Cow = c(1:n_nr_obs),
                                 DIM = vec_dim,
                                 Breed = vec_breed,
                                 Citrate = vec_cit_round[,1])
tbl_cit_dim_br
```


## Plot

```{r}
library(ggplot2)
p <- ggplot(data = tbl_cit_dim_br, 
            aes(x = DIM, y = Citrate)) + 
  geom_point(aes(color = Breed))
p
```


## Write Data
The data is written to a file

```{r}
s_cit_dim_br_path <- here::here("docs", "data", "asm_exam_2024_p02.csv")
# unlink(s_cit_dim_br_path)
if (!file.exists(s_cit_dim_br_path)){
  readr::write_delim(tbl_cit_dim_br, file = s_cit_dim_br_path, delim = ",")
}
```


## Group Means

```{r}
library(dplyr)
tbl_cit_dim_br_gr_mean <- tbl_cit_dim_br %>% 
  group_by(Breed) %>%
  summarise(Cit_Group_Mean = mean(Citrate),
            DIM_Group_Mean = mean(DIM))
tbl_cit_dim_br_gr_mean
```

## Linear Model

```{r}
lm_cit_dim_br <- lm(Citrate ~ DIM + Breed, data = tbl_cit_dim_br)
smry_lm_cit_dim_br <- summary(lm_cit_dim_br)
smry_lm_cit_dim_br
```


Group means

```{r}
smry_lm_cit_dim_br$coefficients["(Intercept)","Estimate"] + 
  tbl_cit_dim_br_gr_mean$DIM_Group_Mean[tbl_cit_dim_br_gr_mean$Breed == "BS"] * 
  smry_lm_cit_dim_br$coefficients["DIM","Estimate"]
```

