---
title: "Repeated Observations"
output: html_notebook
---

## Disclaimer
Some facts on analysis of repeated observations are prepared in this notebook.


## Data
The data we are using is on beef cattle and on blood pressure measurements.

```{r}
s_rep_path_bw_bc <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_bc_rep_obs.csv"
tbl_bw_bc_rep <- readr::read_delim(s_rep_path_bw_bc, delim = ",")
tbl_bw_bc_rep
```

The second data set is on body weight and breed, also with repeated observations.

```{r}
s_rep_path_bw_br <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_breed_rep_obs.csv"
tbl_bw_br_rep <- readr::read_delim(s_rep_path_bw_br, delim = ",")
tbl_bw_br_rep
```


## Plots
The plots of the data already show certain patterns which are important to be considered in any statistical analysis.

```{r}
library(ggplot2)
p_bw_bc_rep <- ggplot(data = tbl_bw_bc_rep,
                      mapping = aes(x = `Breast Circumference`,
                                    y = `Body Weight`)) + 
  geom_point(aes(color = as.factor(Animal)))
p_bw_bc_rep
```

A similar plot with `Breed`

```{r}
library(ggplot2)
p_bw_br_rep <- ggplot(data = tbl_bw_br_rep,
                      mapping = aes(x = as.factor(Breed),
                                    y = `Body Weight`)) + 
  geom_point(aes(color = as.factor(Animal)))
p_bw_br_rep
```

