---
title: "ASMAS SS2024 - Exercise 2"
author: "Peter von Rohr"
date: 2024-02-26
format: html
filters: 
  - webr
packages:
  - readr
  - readxl
  - ggplot2
---

## Problem 1: Reading Data
The first step of a data analysis in R is to read the data. This can be done in different ways which are described below.

### Direct Assignment
As done in Exercise 1, we have assigned the data directly to different R-objects. To recap, this was done with

```{r}
vec_width <- c(82,65,76,80,78,70,72,70,65,73)
vec_height <- c(185,168,168,193,180,181,182,169,165,170)
```


### Reading Files

* From local storage: read data from local file
* From website: specify link directly


### Different Formats

* Excel: An excel file has to be downloaded first and can then be imported.

```{r}
# download first
s_wh_data <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_width_height.xlsx"
s_dest_file <- file.path("~", "Downloads", basename(s_wh_data))
download.file(url = s_wh_data, destfile = s_dest_file)
# read from local file
tbl_wh <- readxl::read_excel(s_dest_file)
tbl_wh
```


* CSV

```{r}
s_wh_data <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_width_height.csv"
tbl_wh <- readr::read_delim(s_wh_data, delim = ",")
tbl_wh
```

The downloaded data can be summarized using the function `summary()`.

```{r}
summary(tbl_wh)
```


## Problem 2: Download Beef-Cattle Data
There is a dataset on `Breast Circumference` and `Body Weight` for beef cattle animals available in two different formats. 

1. Excel: https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_bc_reg.xlsx
2. CSV: https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_bc_reg.csv


### Tasks

* Read the data from both formats
* Provide summary statistics of the variables `Breast Circumference` and `Body Weight`
* Plot `Breast Circumference` on the x-axis and `Body Weight` on the y-axis


### Solutions

* Read the data
Start by reading from Excel workbook

```{webr-r}
# read data from xlsx workbook

```

Read data from CSV-file

```{webr-r}
# read data from CSV-file

```


* Summary statistics for `Breast Circumference` and `Body Weight`

```{webr-r}
# summary statistics

```


* Plot

```{webr-r}
# plot

```

