---
title: "Applied Statistical Methods - Solution 4"
author: "Peter von Rohr"
date: 2024-03-18
filters: 
  - webr
---

```{r}
#| label: asm_sol04_p01_setup
#| include: false
s_data_url_root <- "https://charlotte-ngs.github.io/asmasss2024/data"
s_bw_bc_url <- paste(s_data_url_root, "asm_bw_bc_reg.csv", sep = "/")
```

## Problem 1: Matrix-Vector Notation
Use the dataset on `Body Weight` and `Breast Circumference` to show that the matrix-vector notation of the regression model yields the same result as the scalar notation. 

### Tasks

* Read data from `r s_bw_bc_url`

```{webr-r}
# read data
s_bw_bc_url <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_bc_reg.csv"
df_bw_bc <- read.table(s_bw_bc_url, header = T, sep = ",")
df_bw_bc
```

* Set up the matrix $X$ and the vector $y$

```{webr-r}
# matrix X
X <- matrix(data = c(rep(1, nrow(df_bw_bc)), 
                     df_bw_bc$Breast.Circumference), 
            nrow = nrow(df_bw_bc), ncol = 2)
X
```

the vector $y$ is taken from the dataset directly

```{webr-r}
# vector y
y <- df_bw_bc$Body.Weight
y
```


* Compute the solution for $\hat{b}$

```{webr-r}
# compute solution
xtx <- crossprod(X)
xty <- crossprod(X,y)
b_hat <- solve(xtx,xty)
b_hat
```


## Problem 2: Multiple Linear Regression
Use the dataset on `Body Weight` and multiple other predictor variables to fit a multiple linear regression. 

```{r}
#| label: asm_sol04_p02_setup
#| include: false
s_data_url_root <- "https://charlotte-ngs.github.io/asmasss2024/data"
s_bw_mlr_url <- paste(s_data_url_root, "asm_bw_mult_reg.csv", sep = "/")
```

The dataset is available from `r s_bw_mlr_url`. 

## Tasks
The same as in Problem 1


* Read data from `r s_bw_mlr_url`

```{webr-r}
# read data
s_bw_mlr_url <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_mult_reg.csv"
df_bw_mlr <- read.table(s_bw_mlr_url, header = T, sep = ",")
df_bw_mlr
```

* Set up the matrix $X$ and the vector $y$

```{webr-r}
# matrix X
X <- matrix(data = c(rep(1, nrow(df_bw_mlr)),
                     df_bw_mlr$Breast.Circumference,
                     df_bw_mlr$BCS,
                     df_bw_mlr$HEI),
            nrow = nrow(df_bw_mlr))
X
```

the vector $y$ is taken from the dataset directly

```{webr-r}
# vector y
# vector y
y <- df_bw_mlr$Body.Weight
y
```


* Compute the solution for $\hat{b}$

```{webr-r}
# compute solution
xtx <- crossprod(X)
xty <- crossprod(X,y)
b_hat <- solve(xtx,xty)
b_hat
```


* Validation with `lm()` in R

```{webr-r}
# fit model with lm()
lm_bw_mlr <- lm(Body.Weight ~ Breast.Circumference + BCS + HEI, 
                data = df_bw_mlr)
summary(lm_bw_mlr)
```


