---
title: "Applied Statistical Methods - Solution 6"
author: "Peter von Rohr"
date: 2024-04-15
filters: 
  - webr
---


## Problem 1: Regression On Dummy Variables
```{r, echo=FALSE}
s_ex06p01_data_path <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_flem.csv"
```

Use the dataset with the breeds assigned to every animal and find out the influence of the breed on the response variable `body weight`. The data is available from

```{r echo=FALSE}
cat(s_ex06p01_data_path, "\n")
```

Start by fitting a linear model with `Breed` as the only factor in the model, hence ignore the independent variables such as `Breast Circumference`, `BCS` and `HEI`. 


### Tasks

* Read the data

```{webr-r}
s_ex06p01_data_path <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_flem.csv"
tbl_ex06p01_data <- read.table(s_ex06p01_data_path, 
                               header = T, sep = ",")
tbl_ex06p01_data
```

* Fit a linear model including breed as a factor

```{webr-r}
lm_reg_dummy_bw_breed <- lm(Body.Weight ~ Breed, 
                            data = tbl_ex06p01_data)
summary(lm_reg_dummy_bw_breed)
```
