---
title: "Applied Statistical Methods - Solution 8"
author: "Peter von Rohr"
date: 2024-04-29
filters: 
  - webr
---


# Problem 1: Interactions
Use the following dataset on `Breed`, `Breast.Circumference` and `Body.Weight` and fit a fixed linear effects model with `Body.Weight` as response and `Breed` and `Breast.Circumference` as predictors and include an interaction term between the two predictors. Compute the expected difference in `Body.Weight` for two animals which differ in `Breast.Circumference` by $1cm$ for every `Breed`. 

The dataset is available under

```{r data-ex08-p01, echo=FALSE}
s_tbl_ex08_p01_path <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_flem.csv"
s_tbl_ex08_p01_path
```


## Tasks

* Read the data for fitting the linear model

```{webr-r}
# read data
s_tbl_ex08_p01_path <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_flem.csv"
df_bw_bc_br <- read.table(s_tbl_ex08_p01_path, 
                          header = T, sep = ",")
df_bw_bc_br
```

* Fitting the linear model

```{webr-r}
# lm with interactions
lm_bw_bc_br_int <- lm(Body.Weight ~ Breast.Circumference * Breed, 
                      data = df_bw_bc_br)
# show results with summary
smry_lm <- summary(lm_bw_bc_br_int)
smry_lm
```

* Expected difference in body weight for the three breeds:

__Angus__: The expected difference in body weight (in kg) of one centimeter increase in breast circumference corresponds to the regression coefficient of Breast.Circumference and is 

```{webr-r}
# show estimate for BC
smry_lm$coefficients["Breast.Circumference", "Estimate"]
```

__Limousin__: Because, for the breed limousin, there is an interaction effect. We have to add the regression coefficient of Breast.Circumference to the interaction effect Breast.Circumference:BreedLimousin. From this we get

```{webr-r}
# add slope plus interaction effect for LI
delta_bw_li <- smry_lm$coefficients["Breast.Circumference", "Estimate"] +
  smry_lm$coefficients["Breast.Circumference:BreedLimousin", "Estimate"]
delta_bw_li
```

__Simmental__: The same as for limousin, we have for simmental

```{webr-r}
# add slope plus interaction effect for SI
delta_bw_si <- smry_lm$coefficients["Breast.Circumference", "Estimate"] +
  smry_lm$coefficients["Breast.Circumference:BreedSimmental", "Estimate"]
delta_bw_si
```


## Problem 2: Simulation
```{r, echo=FALSE, message=FALSE, warning=FALSE}
n_b_intercept <- -1070
n_b_slope <- 8.7
n_res_std_error <- 12
n_nr_obs <- 10
n_slope_tol <- 0.1

# take the values for mean and sd for bc from the dataset
s_tbl_bw_bc_br_path <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_flem.csv"
df_bw_bc_br <- read.table(s_tbl_bw_bc_br_path, header = T, sep = ",")
n_mean_bc <- round(mean(df_bw_bc_br$Breast.Circumference), 0)
n_sd_bc <- round(sd(df_bw_bc_br$Breast.Circumference), 2)
```

Use the following values for intercept and regression slope for Body.Weight on Breast.Circumference to simulate a dataset of size $N$. What is the number for $N$ that has to be chosen such that the regression analysis of the simulated data gives the same result as the true regression slope. 

The true values are: 

* Intercept: $`r n_b_intercept`$
* Regression slope: $`r n_b_slope`$
* Residual standard error: $`r n_res_std_error`$

### Hints

* Start with $N=`r n_nr_obs`$, simulate a dataset and analyse the data with `lm()`
* If the result (rounded to 1 digits after decimal point) is not the same then double the size of the dataset, hence use, $N=`r 2*n_nr_obs`$
* Continue until you get close to the true value.
* Assume that the random resiudals follow a normal distribution with mean zero and standard devation equal to $`r n_res_std_error`$
* Take breast circumference to be normally distributed with a mean of $`r n_mean_bc`$ and a standard deviation of $`r n_sd_bc`$
* Use a linear regression model with an intercept to model expected body weight based on breast circumference.


### Tasks

* Assign numbers given in problem description into variables

```{webr-r}
# intercept and slope
n_b_intercept <- -1070
n_b_slope <- 8.7
# residual standard deviation
n_res_std_error <- 12
# number of observations
n_nr_obs <- 10
# mean and sd of BC
n_mean_bc <- 180
n_sd_bc <- 2.6
```

* Start with $N=`r n_nr_obs`$ and first generate the matrix $X$ which consists of a column of all ones and a column of breast circumference values in centimeter taken from the given normal distribution. Whenever, we generate some random numbers it is important to first set the seed with the function `set.seed()` to which an integer number is passed. This makes sure that when repeating the simulation the same results are generated.

```{webr-r}
# start by setting the seed 
set.seed(2904)

vec_bc <- rnorm(n_nr_obs, mean = n_mean_bc, sd = n_sd_bc)
mat_X <- matrix(c(rep(1,n_nr_obs), vec_bc), ncol = 2)
mat_X
```

* Simulate observations of `Body.Weight`

```{webr-r}
# vectors for intercept and slope and a vector for observations
vec_b <- c(n_b_intercept, n_b_slope)
vec_y <- crossprod(t(mat_X), vec_b) + 
  rnorm(n_nr_obs, mean=0, sd=n_res_std_error)
vec_y
```

* Analyse the simulated data with a regression model

```{webr-r}
df_bw_bc_sim <- data.frame(Body.Weight = vec_y, 
                           Breast.Circumference=vec_bc)
lm_bw_bc_sim <- lm(Body.Weight ~ Breast.Circumference, 
                   data = df_bw_bc_sim)
lm_bw_bc_sim
```

* Compute absolute value of deviation between regression and simulation

```{webr-r}
# use function abs()
abs(lm_bw_bc_sim$coefficients[["Breast.Circumference"]] - n_b_slope)
```

* Use a loop to iteratively increase the number of observations until the absolute deviation of the estimated slope from the true value becomes smaller than `r n_slope_tol`.

```{webr-r}
n_slope_tol <- 0.1
n_max_iter <- 10
n_iter_round <- 0
n_abs_dev <- abs(lm_bw_bc_sim$coefficients[["Breast.Circumference"]] - n_b_slope)
while(n_abs_dev > n_slope_tol && 
      n_iter_round < n_max_iter){
  # count number of iterations and 
  # determine number of observations
  n_iter_round <- n_iter_round + 1
  n_nr_obs <- 2 * n_nr_obs
  # simulate breast circumference
  vec_bc <- rnorm(n_nr_obs, mean = n_mean_bc, sd = n_sd_bc)
  mat_X <- matrix(c(rep(1,n_nr_obs), vec_bc), ncol = 2)
  # simulate body weight
  vec_y <- crossprod(t(mat_X), vec_b) + 
    rnorm(n_nr_obs, mean=0, sd=n_res_std_error)
  # analyse simulated data
  df_bw_bc_sim <- data.frame(Body.Weight = vec_y, 
                             Breast.Circumference=vec_bc)
  lm_bw_bc_sim <- lm(Body.Weight ~ Breast.Circumference, 
                     data = df_bw_bc_sim)
  n_abs_dev <- abs(lm_bw_bc_sim$coefficients[["Breast.Circumference"]] - n_b_slope)
  # results
  cat(" * Iteration: ", n_iter_round, "\n")
  cat(" * Number of observations: ", n_nr_obs, "\n")
  cat(" * Regression slope: ", lm_bw_bc_sim$coefficients[["Breast.Circumference"]], "\n")
  
}
```


