---
title: Applied Statistical Methods - `r tools::toTitleCase(params$doctype)` 1
author: Peter von Rohr
date: 2023-02-27
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    keep_md: false
    keep_tex: false
header-includes:
 \usepackage{longtable}
 \usepackage{float}
params:
  doctype:
    label: 'Document Type'
    value: 'solution'
    choices: ['exercise', 'solution', 'notebook']
  isonline:
    label: 'Online (y/n)'
    value: TRUE
    choices: [TRUE, FALSE]
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1: Regression Analysis
The following dataset on body weight and on futher observations on a number of animals is given.

```{r tbl-ex01-p01-data, echo=FALSE, message=FALSE, warning=FALSE}
s_ex01_p01_path <- "https://charlotte-ngs.github.io/asmss2023/data/asm_bw_mult_reg.csv"
tbl_ex01_p01_data <- readr::read_csv(file = s_ex01_p01_path)
knitr::kable(tbl_ex01_p01_data,
             booktabs = TRUE,
             longtable = TRUE,
             escape = FALSE)
```

The data can be read from `r s_ex01_p01_path`. The additional columns contain data on body condition score (BCS) and height (HEI).

### Tasks

* Build a regression model of body weight on the other observations using the dataset given above.
* Set up the matrix $\mathbf{X}$ and the vectors $\mathbf{y}$, $\mathbf{b}$ and $\mathbf{e}$.
* Compute estimate for the regression coefficients in the model defined above.


<!-- your-solution-start
### Your Solution

* Start by building the model which includes the decision of what type of components go into the vector $\mathbf{b}$. Then write down the model in matrix-vector notation.
* Determine all the components of $\mathbf{X}$, $\mathbf{y}$, $\mathbf{b}$ and $\mathbf{e}$.
* Compute the solution for $\hat{\mathbf{b}}

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

First, the data has to be read

```{r, message=FALSE, warning=FALSE}
s_sol01_p01_path <- "https://charlotte-ngs.github.io/asmss2023/data/asm_bw_mult_reg.csv"
tbl_sol01_p01 <- readr::read_csv(file = s_sol01_p01_path)
```

* The regression model can be stated as 

$$ \mathbf{y} = \mathbf{X}\mathbf{b} + \mathbf{e}$$

where 

```{r, echo=FALSE, results='asis'}
n_nr_obs <- nrow(tbl_sol01_p01)
n_nr_pred <- 4
vec_y <- tbl_sol01_p01$`Body Weight`
mat_X <- matrix(c(rep(1, n_nr_obs), 
                  tbl_sol01_p01$`Breast Circumference`, 
                  tbl_sol01_p01$BCS, 
                  tbl_sol01_p01$HEI), ncol = n_nr_pred)
vec_b <- c("b_0", "b_{BC}", "b_{BCS}", "b_{HEI}")
vec_e <- rmdhelp::vecGetVecElem(psBaseElement = "e", pnVecLen = n_nr_obs, psResult = "latex")
cat("$$\n")
cat(paste0(rmdhelp::bmatrix(pmat = mat_X, ps_name = "\\mathbf{X}"), collapse = ""), "\n")
cat("\\text{, }\n")
cat(paste0(rmdhelp::bcolumn_vector(pvec = vec_y, ps_name = "\\mathbf{y}"), collapse = ""), "\n")
cat("\\text{, }\n")
cat(paste0(rmdhelp::bcolumn_vector(pvec=vec_e, ps_name = "\\mathbf{e}"), collapse = ""), "\n")
cat("\\text{, }\n")
cat(paste0(rmdhelp::bcolumn_vector(pvec = vec_b, ps_name = "\\mathbf{b}"), collapse = ""), "\n")
cat("$$\n")
```

The residuals are assumed to have expectation $E(\mathbf{e}) = \mathbf{0}$ and a variance-covariance matrix $var(\mathbf{e}) = I_N * \sigma^2$. 

* The solution for the estimates $\hat{b}$ can be computed as

$$\hat{\mathbf{b}} = (X^TX)^{-1}X^Ty $$

```{r}
mat_xtx <- crossprod(mat_X)
mat_xty <- crossprod(mat_X, vec_y)
mat_sol_bhat <- solve(mat_xtx, mat_xty)
```

The solution vector is

```{r, echo=FALSE, results='asis'}
cat(paste0(rmdhelp::bmatrix(pmat = round(mat_sol_bhat, digits = 3), ps_name = "\\hat{b}", ps_env = "$$"), collapse = ""), "\n")
```


<!-- master-solution-end -->


## Problem 2

Use the same dataset as in Problem 1 and verify your results using the function `lm()` in R. 

<!-- your-solution-start
### Your Solution

* Read the data into a dataframe
* Use the `lm()` function which takes a formula to define the model and a dataframe
* Use the function `summary()` on the result of `lm()` to show the results

---  your-solution-end -->


<!-- master-solution-start -->

### Solution

* Read the data

```{r, message=FALSE, warning=FALSE}
s_sol01_p02_path <- "https://charlotte-ngs.github.io/asmss2023/data/asm_bw_mult_reg.csv"
tbl_sol01_p02 <- readr::read_csv(file = s_sol01_p02_path)
```

* Specify the call to `lm()`

```{r}
lm_bw_mult <- lm(formula = `Body Weight` ~ `Breast Circumference` + `BCS` + `HEI`, 
                 data = tbl_sol01_p02)
summary(lm_bw_mult)
```

* Solutions are obtained by

```{r}
coefficients(lm_bw_mult)
```


<!-- master-solution-end -->
 
