---
title: "ASMAS SS2024 - Exercise 2"
author: "Peter von Rohr"
date: 2024-02-26
format: html
filters: 
  - webr
webr:
  packages: ['readxl']
---

## Problem 1: Reading Data
The first step of a data analysis in R is to read the data. This can be done in different ways which are described below.

### Direct Assignment
As done in Exercise 1, we have assigned the data directly to different R-objects. To recap, this was done with

```{r}
vec_width <- c(82,65,76,80,78,70,72,70,65,73)
vec_height <- c(185,168,168,193,180,181,182,169,165,170)
```


### Reading Files

* From local storage: read data from local file
* From website: specify link directly


### Different Formats

* Excel: An excel file has to be downloaded first and can then be imported.

```{r}
# download first
s_wh_data <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_width_height.xlsx"
s_down_dir <- tempdir()
s_dest_file <- file.path(s_down_dir, basename(s_wh_data))
download.file(url = s_wh_data, destfile = s_dest_file)
# read from local file
tbl_wh <- readxl::read_excel(s_dest_file)
# delete downloaded file
unlink(s_dest_file)
# show table read from xlsx
tbl_wh
```


* CSV

```{r}
s_wh_data <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_width_height.csv"
tbl_wh <- readr::read_delim(s_wh_data, delim = ",")
tbl_wh
```

The downloaded data can be summarized using the function `summary()`.

```{r}
summary(tbl_wh)
```


## Problem 2: Download Beef-Cattle Data
There is a dataset on `Breast Circumference` and `Body Weight` for beef cattle animals available in two different formats. 

1. Excel: https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_bc_reg.xlsx
2. CSV: https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_bc_reg.csv


### Tasks

* Read the data from both formats
* Provide summary statistics of the variables `Breast Circumference` and `Body Weight`
* Plot `Breast Circumference` on the x-axis and `Body Weight` on the y-axis


### Solutions

* Read the data
Start by reading from Excel workbook

```{webr-r}
# read data from xlsx workbook
s_bw_bc_url_xlsx <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_bc_reg.xlsx"
s_down_dir <- tempdir()
s_bw_bc_path <- file.path(s_down_dir, basename(s_bw_bc_url_xlsx))
download.file(s_bw_bc_url_xlsx, s_bw_bc_path)
tbl_bw_bc_xlsx <- readxl::read_excel(s_bw_bc_path)
unlink(s_bw_bc_path)
tbl_bw_bc_xlsx
```

Read data from CSV-file

```{webr-r}
# read data from CSV-file
s_bw_bc_url_csv <- "https://charlotte-ngs.github.io/asmasss2024/data/asm_bw_bc_reg.csv"
df_bw_bc_csv <- read.table(s_bw_bc_url_csv, header = TRUE, sep = ",")
df_bw_bc_csv
```


* Summary statistics for `Breast Circumference` and `Body Weight`

```{webr-r}
# summary statistics
summary(tbl_bw_bc_xlsx)
```


* Plot

```{webr-r}
# plot
plot(tbl_bw_bc_xlsx$`Breast Circumference`, tbl_bw_bc_xlsx$`Body Weight`)
```

